\section{Notes}

----------------------------------------------------------


	Since I am submitting this incredibly late, I will lay out areas I am seeking feedback to make your work easier.
	
	
	For the final draft, I intend to add more robust citations for various parts including citations for the AI interferometry part, as well as for some of the comments I made on accelerometer drift/bias. That is the biggest area I feel this current draft lacks.
	
	I also think I need to add screen shots of the graphics I reference in my text, to make it easier to read/understand what I am talking about.
	
	In particular, the two graphics I really want to include, is the graphics demonstrating the low pass filter which is applied to the inputs of the Kalman Filter, and the final result demonstrating the bias from each of the Kalman Filtered version, non Kalman Filtered version(low pass filtered version with no KF), and the run which doesn't take in to account tidal effects.
	
	As far as feedback I am $looking$ for, I am concerned I don't have enough to say in the discussion or results section. I kind of find that the results are fairly self evident and I am not sure what I have to add. After all the experiment and conclusions are fairly straight forward, but the methods are not. Is there anything I can improve in the explanation sections?
	
	I think I could also add more detail to the AI interferometry section also and dive deeper into the physics, as I spend the most time on the Kalman Filter.
	
	
	
----------------------------------------------------------